---
title: "Apartment Rent Analysis"
output:
  html_document:
    df_print: paged
date: "2022-12-14"
---

```{r, warning=FALSE, message=FALSE}
require(splines)
require(glmnet)
require(viridis)
require(nortest)
require(lmtest)
require(caret)
require(multcomp)
require(broom)
require(tidyverse)
require(mgcv)
require(genlasso)
require(foreach)
require(doParallel)
require(geojsonio)
require(sp)
require(spdep)
require(gratia)
require(visreg)
library(broom)
require(viridis)
require(gridExtra)
set.seed(93)
n.cores <- 1
cluster <- parallel::makeCluster(
  n.cores, 
  type = "FORK"
)
doParallel::registerDoParallel(cl = cluster)
```

# Preparations

The set of apartments with missing square footage filled in via regression from 
the previous stage is loaded. Then, all the categorical variables are converted 
to actual categorical types. Furthermore, the apartments are filtered such that 
only the ones with at most 3 bedrooms, and plausible number of bathrooms compared 
to the bedrooms are retained.


```{r}
apartment_full <- read.csv("full_sqft.csv") %>% filter(beds <= 3) %>% filter((beds == 0 & baths <= 1) | (baths <= beds + 0.5))
apartment_selected <- apartment_full %>% select(-address, -city, -name, -state, -bath_count, -bed_count, -neighborhood)
apartment_selected <- apartment_selected %>% mutate(shopping.num = as.factor(shopping.num),
                                                    zip = as.character(zip),
                                                    count.pub.high = as.factor(count.pub.high), 
                                                    pub.elt.mid = as.factor(pub.elt.mid), 
                                                    priv.elt.mid = as.factor(priv.elt.mid),
                                                    priv.el.hi = as.factor(priv.el.hi),
                                                    priv.elt.mid.hi = as.factor(priv.elt.mid.hi),
                                                    air.level = as.factor(air.level),
                                                    base.num = as.factor(base.num),
                                                    busi.level = as.factor(busi.level),
                                                    has.covered = as.factor(has.covered),
                                                    has.garage = as.factor(has.garage),
                                                    has.lot = as.factor(has.lot),
                                                    has.street = as.factor(has.street),
                                                    pet.allowed = as.factor(pet.allowed),
                                                    traffic.level = as.factor(traffic.level),
                                                    has.parking = as.factor(has.parking),
                                                    has.pub.elementary = as.factor(has.pub.elementary),
                                                    has.priv.elementary = as.factor(has.priv.elementary),
                                                    has.pub.mid = as.factor(has.pub.mid),
                                                    has.cha.high = as.factor(has.cha.high),
                                                    has.priv.high = as.factor(has.priv.high),
                                                    has.cha.mid.high = as.factor(has.cha.mid.high),
                                                    has.priv.mid.high = as.factor(has.priv.mid.high),
                                                    has.pub.mid.high = as.factor(has.pub.mid.high),
                                                    priv.el.hi = as.factor(priv.el.hi),
                                                    cha.elt.mid.hi = as.factor(cha.elt.mid.hi),
                                                    priv.elt.mid.hi = as.factor(priv.elt.mid.hi),
                                                    sqft.regressed = as.factor(sqft.regressed))
```

## Identifying geo-spatial relations between the zip codes

The map of Chicago with zip codes are obtained from the website for the City 
of Chicago. Then, for each zip code, the neighboring zip codes are found, so that 
it would be possible to merge regions with similar effects on rent. Some zip codes 
seem to be split into multiple shapes when just looking at Chicago. Thus, the zips 
with multiple shapes are identified.

```{r}
zip.map <- geojson_read("zip.geojson",  what = "sp")
zip.nb <- poly2nb(zip.map)
zip.mat <- nb2mat(zip.nb, style="B")
zip_freq <- table(zip.map$zip)
duplicates <- names(zip_freq)[which(zip_freq > 1)]
```

Then, prior to getting the final adjacency matrix for zips, the apartment data 
is filtered so that only the apartments in Chicago proper are retained. Then, 
the reference level for zip is set to the one with the most number of apartments, 
so that the reference level actually correspond to a good baseline.

```{r}
apartment_selected <- apartment_selected %>% filter(zip %in% as.character(zip.map$zip))
apartment_selected_zipcount <- aggregate(rent ~ zip, apartment_selected, length)
colnames(apartment_selected_zipcount) <- c("zip", "count")
apartment_selected_zipcount <- apartment_selected_zipcount[order(apartment_selected_zipcount$count, decreasing = T), ]
contrast_zip <- as.character(apartment_selected_zipcount$zip[1])
```

Finally, the adjacency zip matrix is created. The neighboring zip regions for zips 
with duplicated shapes are combined into one. Then, the final matrix is ordered 
in the same fashion as the level of the zip categorical variables from the apartment.

```{r}
options(max.print = 1000000)
dup.zip.idx <- which(zip.map$zip %in% duplicates)
unique.zip.mat <- as.matrix(zip.mat)
distinct_zip_names <- zip.map$zip[-dup.zip.idx]
for(zip.code in duplicates) {
  duplicate_idx <- zip.map$zip == zip.code
  combined_neighbor_row <- apply(unique.zip.mat[duplicate_idx, ], MARGIN = 2, FUN = sum)
  combined_neighbor_row[combined_neighbor_row > 1] <- 1
  combined_neighbor_row <- combined_neighbor_row
  unique.zip.mat <- rbind(unique.zip.mat, combined_neighbor_row)
  unique.zip.mat <- cbind(unique.zip.mat, c(combined_neighbor_row, 0))
  distinct_zip_names <- c(distinct_zip_names, zip.code)
}
unique.zip.mat <- unique.zip.mat[-dup.zip.idx, -dup.zip.idx]
row.names(unique.zip.mat) <- distinct_zip_names
colnames(unique.zip.mat) <- distinct_zip_names

zip_factors <- as.factor(apartment_selected$zip)
zip_factors <- relevel(zip_factors, contrast_zip)
zip.order <- levels(zip_factors)
unique.zip.mat <- unique.zip.mat[zip.order, zip.order]
```

## Variable Selection and Transformations

The original dataset, which came from Apartments.com, had many variables that 
are generated by Apartments.com itself. Those variables include things like nearby 
schools, shopping centers etc. Since those variables are not really well defined, 
and is limited to Apartments.com, they are removed from the dataset. In the end, 
by considering how well defined a variable is, and how easy it is to get a value 
for the variable, the following variables are considered.

- zip: the Zip code
- baths: the number of bathrooms
- beds: the number of bedrooms
- sqft: the size of the apartment in square feet
- sqft.regressed: whether the square feet variable was missing
- has.covered: whether the apartment has covered parking
- has.garage: whether the apartment has parking garages
- has.lot: whether the apartment has uncovered parking lot
- has.street: whether the apartment has street parking

Out of those variables, only the square feet is treated as numerical variable.

```{r}
apartment_selected <- apartment_selected %>% 
  select(sqft, zip, baths, beds, sqft.regressed, has.covered, has.garage, 
         has.lot, has.street, rent) %>% mutate(zip = zip_factors, 
                                               baths = as.factor(baths),
                                               beds = as.factor(beds))

apartment_train <- apartment_selected
```

Since sqft is the only numerical variable, and it is important for the rent, 
transformations on sqft are examined. In the end, a log-log transform seem to 
work well in getting a more linear shape, with less heteroskdacity.

```{r, message=FALSE}
sqft.plt.1 <- ggplot(apartment_train, aes(x = sqft, y = rent)) + geom_point() + geom_smooth(method = "gam")
sqft.plt.2 <- ggplot(apartment_train, aes(x = log(sqft), y = log(rent))) + geom_point() + geom_smooth(method = "gam")
grid.arrange(sqft.plt.1, sqft.plt.2)
```

For the categorical variables, zip, beds, and baths have many levels. Thus, fused 
lasso is used to potentially merge levels together. Fused lasso is a generalization 
of the traditional lasso, where instead of shrinking the coefficients towards zero, 
it shrinks the difference between selected pairs of coefficients, which means that 
the coefficients can be shared for some levels in the categorical variable. Furthermore, 
it can still shrink the coefficients towards 0 with the original penalty term.

For the zip variable, the geolocation adjacency matrix is used to construct the 
penalties. It seems reasonable to assume that nearby zip district may contribute 
similarly to the rent of an apartment. Thus, the coefficients for adjacent zip areas 
are shrinked together.

For the beds and baths variable, a conservative hypothesis would be that adding a 
new beds/baths or half baths would not contribute to more rent. Thus, the coefficients 
for the adjacent count of beds and baths are shrinked together.

In order to economically determine the optimal hyperparameter for lasso, the AIC 
is used to approximate the out of sample error, and the parameter on the lasso path 
corresponding to the least AIC is selected as the optimal tuning parameter.

```{r}
# computes AIC according to generalized LASSO paper
fuse_aic <- function(lasso_path, std_resid) {
    N <- nrow(lasso_path$fit)
    y <- matrix(lasso_path$pathobjs$y, nrow = length(lasso_path$pathobjs$y), ncol = length(lasso_path$lambda), byrow = FALSE)
    error <- apply((y - lasso_path$fit)^2, 2, sum)
    adj.sigma <- N * std_resid^2
    adj.opt <- 2 * std_resid^2 * lasso_path$df
    return((error + adj.sigma + adj.opt) / N)
}

# Utility functions for creating the lasso penalty matrix
set_pen_mat <- function(mat, offset, M) {
  K <- nrow(mat)
  M[(1:K) + offset, (1:K) + offset] <- mat
  return(M)
}

sequential_pen_mat <- function(K, offset, M) {
  mat <- matrix(0, nrow = K, ncol = K)
  for(i in 2:K) {
    mat[i, i - 1] <- 1
    mat[i - 1, i] <- 1
  }
  return(set_pen_mat(mat, offset, M))
}

# Utility function for creating the design matrix and penalty matrix
prep_lasso_mat <- function(apartment_df) {
  # Creates the design matrix with dummy variables
  apartment_train.mat <- as.data.frame(model.matrix(rent ~ . + log(sqft) - sqft, apartment_df)[, -1])
  apartment_train.mat$rent <- apartment_df$rent
  
  # normalizes square feet to between 0 and 1 in order to avoid bias in penalization
  max.logsqft <- max(apartment_train.mat$`log(sqft)`)
  min.logsqft <- min(apartment_train.mat$`log(sqft)`)
  qlogsqft <- (apartment_train.mat$`log(sqft)` - min.logsqft) / (max.logsqft - min.logsqft)
  apartment_train.mat$qlogsqft <- qlogsqft
  apartment_train.mat <- apartment_train.mat %>% select(-`log(sqft)`)
  full.mod.varrank <- lm(log(rent) ~ ., apartment_train.mat)
  M <- length(full.mod.varrank$coefficients)
  pen_mat <- matrix(0, nrow = M, ncol = M)

  # creates penalty for zip by geolocation
  apartment_train.mat <- as.data.frame(model.matrix(rent ~ . - sqft, apartment_df)) %>% mutate(rent = apartment_df$rent)
  zip_coefficients <- full.mod.varrank$coefficients[names(full.mod.varrank$coefficients)[grepl("zip", names(full.mod.varrank$coefficients), fixed = TRUE)]]
  apartment_train_lasso <- data.frame(intercept = apartment_train.mat[, 1])
  apartment_train_lasso <- cbind(apartment_train_lasso, apartment_train.mat[, names(zip_coefficients)])
  zip_numbers <- substring(names(zip_coefficients), 4)
  pen_mat <- set_pen_mat(unique.zip.mat[zip_numbers, zip_numbers], 1, pen_mat)

  # creates penalty for beds sequentially
  beds_coefficients <- full.mod.varrank$coefficients[names(full.mod.varrank$coefficients)[grepl("beds", names(full.mod.varrank$coefficients), fixed = TRUE)]]
  pen_mat <- sequential_pen_mat(length(beds_coefficients), ncol(apartment_train_lasso), pen_mat)
  apartment_train_lasso <- cbind(apartment_train_lasso, apartment_train.mat[, names(beds_coefficients)])
  
  # creates penalty for baths sequentially
  baths_coefficients <- full.mod.varrank$coefficients[names(full.mod.varrank$coefficients)[grepl("baths", names(full.mod.varrank$coefficients), fixed = TRUE)]]
  pen_mat <- sequential_pen_mat(length(baths_coefficients), ncol(apartment_train_lasso), pen_mat)
  apartment_train_lasso <- cbind(apartment_train_lasso, apartment_train.mat[, names(baths_coefficients)])

  # extracts the final columns
  lasso.colnames <- c("intercept", names(zip_coefficients), names(beds_coefficients), names(baths_coefficients))
  colnames(apartment_train_lasso) <- lasso.colnames
  lasso.unincl <- names(full.mod.varrank$coefficients)[!names(full.mod.varrank$coefficients) %in% colnames(apartment_train_lasso)]
  lasso.unincl <- lasso.unincl[-c(1, length(lasso.unincl))]

  apartment_train_lasso <- cbind(apartment_train_lasso, apartment_train.mat[, lasso.unincl])
  apartment_train_lasso$qlogsqft <- qlogsqft
  return(list(X = apartment_train_lasso, penalties = graph_from_adjacency_matrix(pen_mat, mode = "undirected", diag=FALSE)))
}

select_penal_param <- function(apartment_df, max_gamma, n_gamma, sigma, crit = fuse_aic) {
  apartment_fuse_lasso <- prep_lasso_mat(apartment_df)
  gamma_grid <- seq(from = 0, to = max_gamma, length.out = n_gamma)
  best_aic <- Inf
  best_gamma <- 0
  best_lambda <- NULL

  fuse_aic <- function(lasso_path, std_resid) {
    N <- nrow(lasso_path$fit)
    y <- matrix(lasso_path$pathobjs$y, nrow = length(lasso_path$pathobjs$y), ncol = length(lasso_path$lambda), byrow = FALSE)
    error <- apply((y - lasso_path$fit)^2, 2, sum)
    adj.sigma <- N * std_resid^2
    adj.opt <- 2 * std_resid^2 * lasso_path$df
    return((error + adj.sigma + adj.opt) / N)
  }
  
  results <- foreach(
    gamma_param = gamma_grid,
    .combine = "c"
  ) %dopar% {
    gen_lasso_path <- genlasso::fusedlasso(y = log(apartment_df$rent), X = as.matrix(apartment_fuse_lasso$X), graph = apartment_fuse_lasso$penalties, 
                                             maxsteps = Inf, gamma = gamma_param)
    insample_aic <- crit(gen_lasso_path, sigma)
    current_best <- which(insample_aic == min(insample_aic))
    if(length(current_best) >= 2) {
      current_best <- current_best[1]
    }
    finished_gamma = data.frame(completed = c(gamma_param))
    write.csv(finished_gamma, file = "tracker.csv")
    return(list(list(best_aic=insample_aic[current_best], best_idx=current_best, best_lambda = gen_lasso_path$lambda[current_best],  gamma=gamma_param)))
  }
  for(r in results) {
      if(r$best_aic <= best_aic) {
        best_gamma <- r$gamma
        best_aic <- r$best_aic 
        best_lambda <- r$best_lambda
      }
  }
  return(list(aic=best_aic, gamma=best_gamma, lambda=best_lambda))
}
```


After getting the lasso fit from the optimal tuning parameter, the categorical 
variables would need to be consolidated into groups based on the lasso parameters. 


```{r}
combine_lasso_components <- function(matcher, lasso_betas, zero_level) {
  lasso_betas <- sort(lasso_betas)
  coef_names <- names(lasso_betas)
  coef_names_match <- coef_names[grepl(matcher, coef_names, fixed = TRUE)]
  betas_match <- lasso_betas[coef_names_match]
  betas_match <- sort(betas_match)
  combine_mat <- matrix(0, nrow = length(betas_match), ncol = length(betas_match))
  
  for(i in 2:length(betas_match)) {
    if(betas_match[i] - betas_match[i - 1] == 0) {
      combine_mat[i, i - 1] = 1
      combine_mat[i - 1, i] = 1
    }
  }
  combine_graph <- graph_from_adjacency_matrix(combine_mat, mode = "undirected", diag=FALSE)
  combine_components <- components(combine_graph, "weak")
  select_components <- which(combine_components$csize >= 2)
  result = list()
  zero_components <- c()
  zip_factors
  for(c in select_components) {
    component_parts <- names(betas_match[combine_components$membership == c])
    if(sum(abs(betas_match[component_parts])) != 0) {
      result[[paste0("C", c)]] <- component_parts
    }
  }
  
  zero_components <- c(names(betas_match[betas_match == 0]), zero_level)
  if(length(zero_components) > 1) {
    result[["C0"]] <- zero_components
  }
  
  return(result)
}

consolidate_levels <- function(partitions, prefix_length, var, df) {
  zero_levels <- levels(df[[var]])[1]
  df <- df %>% mutate(!! var := as.character(.data[[var]]))
  for(p in names(partitions)) {
    var_level <- substring(partitions[[p]], prefix_length)
    df <- df %>% 
      mutate(!! var := case_when(.data[[var]] %in% var_level ~ p, TRUE ~ .data[[var]]))
  }
  df <- df %>% mutate(!! var := as.factor(.data[[var]]))
  if("C0" %in% names(partitions)) {
    df[[var]] <- relevel(df[[var]], "C0")
  } else {
    df[[var]] <- relevel(df[[var]], zero_levels)
  }
  return(df)
}

apartment_train <- apartment_train %>% mutate(beds = factor(beds, levels = sort(unique(beds)))) %>% mutate(baths = factor(baths, levels = sort(unique(baths))))
apartment_merged <- as.data.frame(apartment_train)
```

## Model Selection and 10-Fold CV

Since there will not be any more apartment from May of 2022, the model is fitted 
mostly to find the relationships between the variables. Thus, only cross validation 
well be performed. For cross validation, the dataset is shuffled then split into 
10 folds. For each fold, the lasso consolidation is performed first. Then, in order to 
determine the appropriate level of complexity and transformation on the square feet, 
three ways of handling the square feet/rent are attempted. First, a log-log linear model 
with thin-plate smooth on square feet is tried. Finally, a log log GAM model thin-plate 
smooth on log(sqft) is tested as well. Finally, the models 
are tested on the fold, and the average test RMSE is calculated.

```{r}
apartment_train_shuffled <- apartment_train[sample(1:nrow(apartment_train)), ]
apartment_folds <- groupKFold(1:nrow(apartment_train_shuffled), 10)

cv_rmse <- data.frame()
for(indexes in apartment_folds) {
  training_df <- apartment_train[indexes, ]
  full.mod.log.2 <- lm(log(rent) ~ . - sqft + log(sqft), training_df)
  best_lasso <- select_penal_param(training_df, 1, 200, sd(full.mod.log.2$residuals))
  apartment_fuse_lasso <- prep_lasso_mat(training_df)
  gen_lasso_path <- genlasso::fusedlasso(y = log(training_df$rent), X = as.matrix(apartment_fuse_lasso$X), graph = apartment_fuse_lasso$penalties, 
                                             maxsteps = Inf, gamma = best_lasso$gamma)
  fuse_lasso_coef <- coef(gen_lasso_path, lambda = best_lasso$lambda)
  lasso_betas <- c(fuse_lasso_coef$beta)
  names(lasso_betas) <- colnames(apartment_fuse_lasso$X)
  apartment_merged <- as.data.frame(apartment_train)
  zip_partitions <- combine_lasso_components("zip", lasso_betas, paste0("zip", levels(apartment_merged$zip)[1]))
  baths_partitions <- combine_lasso_components("baths", lasso_betas, "baths1")
  beds_partitions <- combine_lasso_components("beds", lasso_betas, "beds0")
  apartment_merged <- consolidate_levels(zip_partitions, 4, "zip", apartment_merged)
  apartment_merged <- consolidate_levels(baths_partitions, 6, "baths", apartment_merged)
  apartment_merged <- consolidate_levels(beds_partitions, 5, "beds", apartment_merged)
  training_merged <- apartment_merged[indexes, ]
  test_df <- apartment_merged[-indexes, ]
  
  log.log.mod <- lm(log(rent) ~ log(sqft) + zip + baths + beds + sqft.regressed + has.covered + has.garage + has.lot + has.street, data = training_merged)
  untrans.gam <- gam(rent ~ s(sqft, k = 2001, m = 2) + zip + baths + beds + sqft.regressed + has.covered + has.garage + has.lot + has.street, data = training_merged, method = "REML")
  gam.mod.log <- gam(log(rent) ~ s(log(sqft), k = 2001, m = 2) + zip + baths + beds + sqft.regressed + has.covered + has.garage + has.lot + has.street, data = training_merged, method = "REML")
  
  log.log.pred <- exp(predict(log.log.mod, test_df))
  untrans.pred <- predict(untrans.gam, test_df)
  gam.log.pred <- exp(predict(gam.mod.log, test_df))
  
  log.log.pred.rmse <- sqrt(mean((test_df$rent - log.log.pred)^2))
  untrans.pred.rmse <- sqrt(mean((test_df$rent - untrans.pred)^2))
  gam.pred.rmse <- sqrt(mean((test_df$rent - gam.log.pred)^2))
  
  cv_rmse <- rbind(cv_rmse, c(log.log.pred.rmse, "log.log.lm"))
  cv_rmse <- rbind(cv_rmse, c(untrans.pred.rmse, "reg.gam"))
  cv_rmse <- rbind(cv_rmse, c(gam.pred.rmse, "log.gam"))
}
apartment_merged <- apartment_merged %>% mutate(zip = as.factor(zip))
levels(apartment_merged$zip)
colnames(cv_rmse) <- c("RMSE", "Method")
cv_rmse$RMSE <- as.numeric(cv_rmse$RMSE)
head(cv_rmse)
```

From the box plot of 10-fold test performances, it seems that the log log GAM 
model had the lowest shifted IQR range. Thus, the log log GAM model is picked as the 
final model to fit.

```{r}
ggplot(cv_rmse, aes(x = Method, y = RMSE, fill = Method)) + geom_boxplot() + 
  ggtitle("Cross Validation Performance")
```

Similar to the steps taken in cross validation, the fused LASSO is applied to 
consolidate the categorical variables. Then, the GAM model with smoothing applied 
to the logged square feet is fitted to predict logged rent.

```{r}
best_lasso <- select_penal_param(apartment_train, 1, 201, 0)
apartment_fuse_lasso <- prep_lasso_mat(apartment_train)
gen_lasso_path <- genlasso::fusedlasso(y = log(apartment_train$rent), X = as.matrix(apartment_fuse_lasso$X), graph = apartment_fuse_lasso$penalties, 
                                             maxsteps = Inf, gamma = best_lasso$gamma)
fuse_lasso_coef <- coef(gen_lasso_path, lambda = best_lasso$lambda)
lasso_betas <- c(fuse_lasso_coef$beta)
names(lasso_betas) <- colnames(apartment_fuse_lasso$X)
apartment_merged <- as.data.frame(apartment_train)
zip_partitions <- combine_lasso_components("zip", lasso_betas, paste0("zip", levels(apartment_merged$zip)[1]))
baths_partitions <- combine_lasso_components("baths", lasso_betas, "baths1")
beds_partitions <- combine_lasso_components("beds", lasso_betas, "beds0")
apartment_merged <- consolidate_levels(zip_partitions, 4, "zip", apartment_merged)
apartment_merged <- consolidate_levels(baths_partitions, 6, "baths", apartment_merged)
apartment_merged <- consolidate_levels(beds_partitions, 5, "beds", apartment_merged)
gam.mod.log.2 <- gam(log(rent) ~ s(log(sqft), k = 2001, m = 2) + zip + baths + beds + sqft.regressed + has.covered + has.garage + has.lot + has.street, data = apartment_merged, method = "REML")
```

While beds and baths are untouched, two zip codes are consolidated together via fused LASSO.

```{r}
zip_partitions
```

## Model Assessment

The summary for the model is given below. The model was able to achieve a $R^2$ of 
around 79% with an adjusted $R^2$ of 78.4%. Although the distributional assumptions 
are not checked, almost all of the terms are very significant with p-value almost 
0. 

```{r}
summary(gam.mod.log.2)
```

The normal QQ plot of the residuals is shown below. It can be seen that the residuals start 
to deviate from the reference line after the theoretical quantiles exceed 2 or -2. Thus, 
for the bulk of the observations, the residuals are generally falling along a normal 
distribution.

```{r}
residuals_df <- data.frame(
  Observed = apartment_merged$rent,
  Predicted = gam.mod.log.2$fitted.values
)
ggplot(residuals_df, aes(sample = Predicted)) + stat_qq() + stat_qq_line() + xlab("Theoretical Quantiles") + 
  ylab("Observed Quantiles") + ggtitle("Residual QQ-Plot")
```

However, due to the large sample size, and the fact that the residuals do deviate from 
a normal distribution eventually, the lilliefour's test for normality came out to 
have strong evidence for the residuals to be not norma.

```{r}
lillie.test(gam.mod.log.2$residuals)
```

An examination of the distribution of the residuals at various points shows that as 
the logged observed rent got larger, the estimator goes from over-estimating the 
log rent to under-estimating the rent. Thus, it suggests that the model is 
biased with other factors potentially not accounted for.

```{r}
residuals_df <- residuals_df %>% mutate(Binned = floor(log(Observed) / 0.1) * 0.1) %>% 
  mutate(Binned = as.factor(Binned)) %>% mutate(Residual = log(Observed) - Predicted)
ggplot(residuals_df, aes(x = Binned, y = Residual)) + geom_boxplot() + ggtitle("Distribution of Residuals")
```

While the box plots seems to have similar widths, a BP test shows that there are 
strong evidence against the hypothesis that the variance of the residuals are 
equal.

```{r}
bptest(gam.mod.log.2)
```

Finally, a DW test for auto-correlation on the residuals show that there are significant 
auto-correlation in the model. Thus, the residuals are not independent of each other.

```{r}
dwtest(gam.mod.log.2, alternative = c("two.sided"))
```

The failure of the assumptions indicates that a more careful and conservative interpretation of the 
significance of the terms should be adopted. However, since a vast majority of terms are 
highly significant, and the tests/confidence intervals for least square regression are 
usually robust even with some violation of the assumptions, it seems that the model 
can still reveal much notable factors that contributes to the rent.

## Model Interpretation

### Square Feet

The multiplicative effect of square feet is plotted below for all range of square feet. 
It can be seen that as the data gets thinner as the square footage increases, the spline 
became more unstable with increasing width of confidence intervals. Towards the 
5000 square feet apartment, where there seems to be only a few data points, the 
confidence interval is from a bit less than 2.5 to almost 10 times. Thus, for the 
high square feet apartments, the model would not be able to give much insight, besides 
that their square footage contributed positively towards their rent, due to the lack 
of the data points.

```{r}
get_effect <- function(l_rent) {
  return(exp(l_rent - coef(gam.mod.log.2)[1]))
}

visreg(gam.mod.log.2, "sqft", trans = get_effect, cond=list(zip="60601", beds = "0", baths = "1", 
                                sqft.regressed = "False", has.covered = "False",
                                has.garage = "False", has.lot = "False", has.street = "False"), 
       ylab = "Multiplicative Effect", main = "Multiplicative Effect of Square Feet")
```

Now, restricting the attention to the segment with more data, which would be the 
apartments with less than 2000 sqft, there is a general increasing effect with greater 
square footage. Furthermore, as the square feet increases, compared to the smaller 
apartments, the increase in multiplicative effect for the larger apartment is 
diminishing. For a small group of dorm sized apartments, the small size lead to 
a reduction in the rent. However, the effect quickly became positive as the 
square footage approaches 500 sqft. Finally, it is worth noting that the the effect 
curve is not very smooth. Combined with the fact that there may be omitted effects 
that lead to bias from the model diagnosis earlier, it may be that the square feet 
have acted as a proxy variable for some unincluded factors in the model. For instance, 
as the apartments get larger, they are likely to become more luxurious with more amenities. 
Then, for the larger apartments, there are several spikes in the estimated effects of 
square feet. It is possible that the spikes may be actually better explained by factors 
such as amenities, which are not included. Thus, a potential improvement for the model 
is to reliably include more factors, which could make it more clear on whether the square 
footage has an actually complicated structure with respect to effects on rent.

```{r}
sqft_full_vs <- visreg(gam.mod.log.2, "sqft", trans = get_effect, 
       data = apartment_merged, ylab = "rent", 
       cond=list(zip="60601", beds = "1", baths = "1", 
                                sqft.regressed = "False", has.covered = "False",
                                has.garage = "False", has.lot = "False", has.street = "False"), 
       plot = FALSE)
sqft_sub_vs <- subset(sqft_full_vs, sqft <= 2000)
plot(sqft_sub_vs, ylab = "Multplicative Effect", main = "Multiplicative Effect of Square Feet")
```

### Zip Codes

The multiplicative effects of zip codes are added to the geospatial data of 
Chicago zip codes in order to visualize the effects of zip codes.

```{r}
gam.mod.coef <- coef(gam.mod.log.2)
zip_coefs <- gam.mod.coef[grepl("zip", names(gam.mod.coef), fixed = TRUE)]
zip_coef_df <- data.frame()

for(n in names(zip_coefs)) {
  zipcode <- substring(n, 4)
  if(zipcode %in% names(zip_partitions)) {
    actual_codes <- substring(zip_partitions[[zipcode]], 4)
    for(code in actual_codes) {
      zip_coef_df <- rbind(zip_coef_df, c(code, zip_coefs[n]))
    }
  } else {
    zip_coef_df <- rbind(zip_coef_df, c(zipcode, zip_coefs[n]))
  }
}
if(!grepl("C0", levels(apartment_merged$zip)[1], fixed = T)) {
  zip_coef_df <- rbind(zip_coef_df, c(levels(apartment_merged$zip)[1], 0))
}

colnames(zip_coef_df) <- c("zip", "effect")
zip_coef_df$effect <- as.numeric(zip_coef_df$effect)
zip_id_df <- data.frame(id = zip.map$objectid, zip = zip.map$zip)
zip_coef_df <- merge(x = zip_id_df, y = zip_coef_df, by = "zip", all.x = TRUE)
zip.geo.df <- tidy(zip.map)
zip.geo.df <- merge(x = zip.geo.df, y = zip_coef_df, by = "id", all.x = TRUE)
zip.geo.df$effect <- exp(zip.geo.df$effect)
```

The map of multiplicative effects by zip codes is shown below. As mentioned before, 
the reference level for zip is zip area with the median rent. From the map, it seems 
that the coastal areas, as well as the areas around and north of downtown correspond 
to higher apartment rent. Then, the southern area inland as well as the most northern 
zip codes have a reduced rent compared to the median zip code. While the relatively lower 
rent associated with the southern area is understandable due to the security issue in 
south Chicago, it is surprising that the northern tip also has a reduced rent. Some 
Google search reveals that it may be entering Northern Chicago, which is not the same 
as Chicago. Consequently, it may be transitioning into lower density areas with land use 
not as tight as Chicago proper.

```{r}
ggplot(data = zip.geo.df) +
  geom_polygon(aes( x = long, y = lat, group = group, fill=effect), color="white") +
  theme_void() + scale_fill_viridis() +
  coord_map() + ggtitle("Multiplicative Effect to Rent by Zip")
```

### Beds and Baths

The confidence intervals and point estimates for beds and baths are extracted 
from the summary information of the model.

```{r}
confint.paragam <- function(mod, level = 0.95) {
  mod.s <- summary(mod)
  
  E <- data.frame(Estimate = mod.s$p.coeff) %>%
    mutate(Term = row.names(.)) %>%
    select(Term, Estimate)
  
  SE <- data.frame(SE = mod.s$se) %>%
    mutate(Term = row.names(.)) %>%
    select(Term, SE)
  
  nu <- mod.s$residual.df
  
  inner_join(E, SE) %>%
    mutate(L = Estimate +
             SE * qt(df = nu,
                     p = (1 - level) / 2),
           U = Estimate +
             SE * qt(df = nu,
                     p = 1 - (1 - level) / 2)) %>%
    return
  
}

gam.confint <- confint.paragam(gam.mod.log.2) %>% select(Term, Estimate, L, U)
```

There are significant multiplicative effects on the rent depending on whether it's a 
studio, or apartment with various number of bedrooms. The increase in rent from 
a studio to a single bedroom apartment is slightly lower than 1.15 times. Then, for each 
subsequent number of bedrooms, the apartments get around 0.05 times more expensive compared 
to the studio apartments.

```{r}
gam.confint.beds <- gam.confint %>% filter(grepl("beds", Term, fixed = T))
ggplot(gam.confint.beds, aes(x = Term, y = exp(Estimate), color = Term)) + scale_y_continuous(limits = c(1, 1.4), breaks = seq(1, 1.4, 0.05)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin=exp(L), ymax=exp(U)), width=.3) + xlab("Bedrooms") + ylab("Multiplicative Effect") + 
  ggtitle("Multiplicative Effect of Bedrooms") 
```

A different way of looking at the contribution of bedrooms is to consider the increase 
in rent for an additional bedroom. Then, despite the increase in rent compared to 
the studio case being almost linear in the previous plot, the marginal increase 
for each bedroom is diminishing. While going from studio to one bedroom would lead 
to an approximately 13% increase in rent, going from one bedroom to two would only 
lead to an approximately 7% increase in rent. There does not seem to be significant 
difference in the marginal increase from two to three bedroom compared to one to two. 
Considering that as the bedroom increases, the rent increases, so that a multiplicative 
increase from the previous number of bedrooms would have a greater increase in raw rent, 
the decrease in marginal multiplicative effect can control the actual raw increase 
in rent.

```{r}
beds.marginal.confint <- confint(glht(gam.mod.log.2, linfct = c("beds2 - beds1 = 0", "beds3 - beds2 = 0")))
beds.marginal.df <- as.data.frame(beds.marginal.confint$confint)
colnames(beds.marginal.df) <- names(gam.confint.beds[1, 2:4])
beds.marginal.df <- rbind(gam.confint.beds[1, 2:4], beds.marginal.df)
row.names(beds.marginal.df) <- 1:3
ggplot(beds.marginal.df, aes(x = 1:3, y = exp(Estimate))) + scale_x_continuous(limits = c(0.8, 3.2), breaks = c(1, 2, 3)) +
  geom_point(size = 3) + geom_line() + scale_y_continuous(limits = c(1, 1.16), breaks = seq(1, 1.15, 0.05)) +
  geom_errorbar(aes(ymin=exp(L), ymax=exp(U)), width=.3) + xlab("Bedrooms") + ylab("Marginal Increase") + 
  ggtitle("Marginal Effect of Bedrooms") 
```


The bathroom case seems to be more interesting. As the number of bathroom increases, 
there seems to be more variety in the rent of the apartments. The addition of a 
half-bath can be attributed with around 8% increase in the rent from a single bathroom 
apartment. Then, adding a full-bathroom from a single bathroom can be attributed with around 
20% increase from a single bathroom. Interestingly, 2 bathrooms with a half bath 
is not significantly different from 3 bathrooms, where both can be attributed with around 
35-40% increase in rent from single bath. Finally, having 3 full baths with a half bath 
can be attributed with an increase in rent on the magnitude of 55-100% compared to 
single bath. The large confidence interval can be attributed to the relatively small 
number of apartment with these many bathrooms. However, it is clear that despite the 
large range, apartments with 3 bath and a half bath are fairly luxurious compared 
to the single bath apartments. In fact, even starting from the 2.5 bathroom apartments, 
the increase in multiplicative effect is disproportional to the increase seen in 
the 1.5 and 2 bath apartments. The disproportional change may suggests that the bathrooms 
are correlated with some form of luxuriousness as the apartments become sort of like a 
house.

```{r}
gam.confint.baths <- gam.confint %>% filter(grepl("baths", Term, fixed = T))
ggplot(gam.confint.baths, aes(x = Term, y = exp(Estimate), color = Term)) + 
  geom_point(size = 3) + geom_errorbar(aes(ymin=exp(L), ymax=exp(U)), width=.3) +
  scale_y_continuous(limits = c(1, 2.05), breaks = seq(1, 2.05, 0.1)) +
  xlab("Bathrooms") + ylab("Multiplicative Effect") + 
  ggtitle("Multiplicative Effect of Bathrooms") 
```

The point marginal effect of bathrooms is increasing with the exception of three bathrooms 
vs 2.5 bathrooms. However, there are some overlaps in the confidence intervals, so that 
the difference in marginal effects are not as significant as the bedroom case. Nevertheless, 
intuitively, it would seem unlikely that for each additional bathroom/half-bath, one has to pay 
an increasing extra amount of rent just because of the added bathroom. It seems more plausible 
that the number of bathroom is a proxy for other factors, such as the apartment having 
two floors for instance, where a half bath would be located on the first floor.

```{r}
baths.marginal.confint <- confint(glht(gam.mod.log.2, linfct = c("baths2 - baths1.5 = 0", 
                                                                 "baths2.5 - baths2 = 0",
                                                                 "baths3 - baths2.5 = 0",
                                                                 "baths3.5 - baths3 = 0")))
baths.marginal.df <- as.data.frame(baths.marginal.confint$confint)
colnames(baths.marginal.df) <- names(gam.confint.baths[1, 2:4])
baths.marginal.df <- rbind(gam.confint.baths[1, 2:4], baths.marginal.df)

ggplot(baths.marginal.df, aes(x = seq(1.5, 3.5, 0.5), y = exp(Estimate))) +
  geom_point(size = 3) + geom_line() + scale_x_continuous(limits = c(1.3, 3.7), breaks = seq(1.5, 3.5, 0.5)) +
  geom_errorbar(aes(ymin=exp(L), ymax=exp(U)), width=.3) + xlab("Bathrooms") + ylab("Marginal Increase") + 
  ggtitle("Marginal Effect of Bathrooms") 
```

